Read Me

### 2023-04-21 code/processing_code/1_cleaning_2020.Rmd
added more to project
more stuff

Load the files to be cleaned.
```{r}
deaths <- read_xlsx(here("data/raw_data/AIDSVu_State_Death_2020.xlsx"), skip = 3)
us_race_pop <- read_xlsx(here("data/raw_data/DECENNIALPL2020.P2-2023-02-24T124235.xlsx"), sheet = 2)
us_sex_pop <-read_csv(here("data/raw_data/raw_data.csv"), skip = 2)
kid_pop <- read_xlsx(here("data/raw_data/Child population by age group.xlsx"))
kid_race <- read_xlsx(here("data/raw_data/Child population by race and ethnicity.xlsx"))

new_dx <- read_xlsx(here("data/raw_data/AIDSVu_State_NewDX_2020.xlsx"), skip = 3)
late_dx <- read_xlsx(here("data/raw_data/AIDSVu_State_LateDiagnoses_2020.xlsx"), skip = 3)
prev  <- read_xlsx(here("data/raw_data/AIDSVu_State_Prev_2020.xlsx"), skip = 3)
rec_care  <- read_xlsx(here("data/raw_data/AIDSVu_State_ReceiptofCare_2020.xlsx"), skip = 3)
sdoh  <- read_xlsx(here("data/raw_data/AIDSVu_State_SDOH_2020.xlsx"), skip = 3)
v_sup  <- read_xlsx(here("data/raw_data/AIDSVu_State_ViralSuppression_2020.xlsx"), skip = 3)
```

@@

summary(deaths_race)
```
Some rates are -9 due to missingness. Let's convert these to NA
```{r}
deaths_race <- deaths_race %>% 
  mutate(across(ends_with("Death Rate"), ~ case_when( .x < 0 ~ NA, TRUE ~ .x)))
```

@@

Some quick cleaning of the social determinant sof health file, and merging this with our two results dataframes `sex_df` and `race_df`.
```{r}
sdoh <- sdoh %>% 
  filter(!grepl("DC", `State Abbreviation`)) %>% 
   filter(!grepl("PR", `State Abbreviation`))

@@

@@

```{r}
rm(deaths, df_pop, kid_pop, pop_sex, pop_less13, us_sex_pop, us_race_pop, new_pop, deaths_sex, race_pop, kid_race, total_pop, deaths_race, late_dx, new_dx, sex_dx, prev, prev_sex, prev_race, v_sup, rec_care, sdoh)
```

>>>>

Load the files to be cleaned.
```{r}
deaths <- read_xlsx(here("data/raw_data/AIDSVu_State_Death_2020.xlsx"), skip = 3)
us_race_pop <- read_xlsx(here("data/raw_data/DECENNIALPL2020.P2-2023-02-24T124235.xlsx"), sheet = 2)
us_sex_pop <-read_csv(here("data/raw_data/raw_data.csv"), skip = 2) # not used in this analysis
kid_pop <- read_xlsx(here("data/raw_data/Child population by age group.xlsx"))
kid_race <- read_xlsx(here("data/raw_data/Child population by race and ethnicity.xlsx"))

new_dx <- read_xlsx(here("data/raw_data/AIDSVu_State_NewDX_2020.xlsx"), skip = 3)
late_dx <- read_xlsx(here("data/raw_data/AIDSVu_State_LateDiagnoses_2020.xlsx"), skip = 3)
prev  <- read_xlsx(here("data/raw_data/AIDSVu_State_Prev_2020.xlsx"), skip = 3)
rec_care  <- read_xlsx(here("data/raw_data/AIDSVu_State_ReceiptofCare_2020.xlsx"), skip = 3)
sdoh  <- read_xlsx(here("data/raw_data/AIDSVu_State_SDOH_2020.xlsx"), skip = 3)
v_sup  <- read_xlsx(here("data/raw_data/AIDSVu_State_ViralSuppression_2020.xlsx"), skip = 3)
internet <- read_xlsx(here("data/raw_data/ACSST5Y2020.S2801-2023-04-17T232857.xlsx"), skip = 3, sheet = 2)
```

@@

summary(deaths_race)
```

Some rates are -9 due to missingness. Let's convert these to NA
```{r}
deaths_race <- deaths_race %>% 
  mutate(across(ends_with("Death Rate"), ~ case_when( .x < 0 ~ NA, TRUE ~ .x)))
```

@@

Some quick cleaning of the social determinants of health file, and merging this with our two results dataframes `sex_df` and `race_df`.
```{r}
sdoh <- sdoh %>% 
  filter(!grepl("DC", `State Abbreviation`)) %>% 
   filter(!grepl("PR", `State Abbreviation`))

@@

```{r}
internet <- internet %>% 
  select(c(1, "Broadband of any type")) %>% 
  slice(c(-1:-7, -48:-52,-263:-267)) %>% 
  filter(!grepl(",", `Broadband of any type`))
```

This dataframe has some issues. Let's fill in the state names...
```{r}
# create a vector of state names
states <- state.name

# repeat each state name five times
states <- states %>%
  rep(each = 4)

# sort the repeated state names in alphabetical order
states <- sort(states)

# print the resulting vector
states

internet$State<- states

internet <- internet %>% 
  select(State, `Broadband of any type`) %>% 
  drop_na()

internet$`Broadband of any type` <- gsub("%", "", internet$`Broadband of any type`)

internet <- internet %>% mutate(`Broadband of any type` = as.numeric(`Broadband of any type`))
```

```{r}
race_df <- full_join(race_df, internet)
sex_df <- full_join(sex_df, internet)
```

@@

### 2023-04-21 code/processing_code/2b_cleaning_race_df.Rmd
more stuff

```{r}
black_df <- race_df %>% select(1:3, contains("Black"), 50:62)

@@

hispanic_df <- race_df %>% select(1:3, contains("Hispanic"), 50:62)

@@

white_df <- race_df %>% select(1:3, contains("White"), 50:62)

@@

multiple_df <- race_df %>% select(1:3, contains("Multiple"), 50:62)

>>>>

```{r}
black_df <- race_df %>% select(1:3, contains("Black"), 50:63)

@@

hispanic_df <- race_df %>% select(1:3, contains("Hispanic"), 50:63)

@@

white_df <- race_df %>% select(1:3, contains("White"), 50:63)

@@

multiple_df <- race_df %>% select(1:3, contains("Multiple"), 50:63)
